---
title: 'Research Studio: Visualizing Data in R'
author: "Cindy Traub, PhD"
date: "September 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("ggplot2")
library(ggplot2)
```

# What would you like to plot today?

R is a fantastic resource for cleaning, aggregating, and analyzing data of many types. Our goal in this Research Studio is to introduce several types of plots you can easily produce from a dataframe within R.

## Sample data included in [ggplot2](http://docs.ggplot2.org/current/#data)

We look at the [diamonds](http://docs.ggplot2.org/current/diamonds.html) data set. Follow the link to access the metadata.

# Some common plots (using base R)

We can examine the diamonds data in several text-based ways:

```{r diamonds}
head(diamonds)
class(diamonds)
str(diamonds)
summary(diamonds)
names(diamonds)
colSums(is.na(diamonds))
table(diamonds$color)
table(diamonds$cut)
```

Let's look at some basic visuals using this data:

```{r diamondbaseplots}
plot(diamonds$color)
plot(diamonds$price)
plot(sort(diamonds$price))
plot(diamonds$carat, diamonds$price)
hist(diamonds$price)
boxplot(diamonds$price ~ diamonds$clarity)
```

We can add lines, labels, etc. But if you want to change color, position, or other features, you have to run each of the three lines again.

```{r caratPriceLabels}
plot(diamonds$carat, diamonds$price)
abline(h=8000, col="red")
text(4, 9000, "Two months salary", col="red")
plot(sort(diamonds$carat))
abline(h=.5, col="blue")
abline(h=.75, col="green")
abline(h=1, col="red")
abline(h=1.5, col="papayawhip")
```

## How to search the help

Type ? followed by your command of interest, for example ?abline or ?plot.

### Task for you: use the help file to determine valid color names that R would know.

The help files can be a bit daunting at first. You often don't need to read them from start to end like a novel. They have a standardized structure, and will give you detail on parameters you can set for the different functions you use

# Some common plots (using ggplot)

Note that we can store a collection of the parameters to a variable, then call that to display the plot. This is useful when you want to repeat some features of your setup.

```{r diamond_ggplots}
qplot(x=carat, y=price, data=diamonds, geom="point")
g<-ggplot(data=diamonds, aes(x=carat, y=price)) 
g + geom_point()
g + geom_point() + ggtitle("The impact of diamond size on price")
g + geom_point() + geom_abline(slope=0, intercept = 8000, col="red") +ggtitle("Scatter of carat vs price")
```

## Faceting

Often it is useful to visualize subsets of your data.

```{r facet_diamonds}
g + geom_point() + ggtitle("The impact of diamond size on price") + facet_grid(color ~clarity, labeller = label_both)
```

Perhaps coloring by one and faceting by the other is more informative:

```{r facet2_diamonds}
g_faceted<-ggplot(data=diamonds, aes(x=carat, y=price, color=color)) + geom_point(alpha=.1) + ggtitle("The impact of diamond size on price") + facet_grid(. ~clarity, labeller = label_both)
g_faceted
```

# Exporting your plots

```{r export}
ggsave("plot.png", width=5, height=5)
```
